Merging merges, more or less

### 2015-04-16 2015mergingmerges/section/02-Machines.tex
change some @move@ rules over to new style

>>>>

\subsection{Definitions}
\TODO{define some interesting functions:}

$@channel@~t$ takes a state type (eg $@Out@~f~n$) and returns the channel it is operating on ($n$), if applicable.
States like @Skip@ and @Done@ don't have an associated channel.

$@inputs@~m$ is the set of all channels used in @Pull@s, @Release@s and @Close@s.

$@outputs@~m$ is the set of all channels used in @Out@s, @OutDone@s, @Update@s and @If@s.
While @Update@s and @If@s do not perform outputs, they must belong to a particular output channel, and including them in @outputs@ simplifies handling of degenerate cases where @Update@s are performed but no @Out@s.

$@initial@~m$ gives the initial state of the machine.

$@states@~m$ gives the set of states of the machine, ignoring transitions.

### 2015-04-16 2015mergingmerges/section/03-Merging.tex
change some @move@ rules over to new style

\begin{tabbing}
MMMMMMM \= MM \= MMMMMM \= M\kill
@merge@ \> $:$ \> $@Machine @l_1 \times @Machine @l_2$ \\
        \> $\to$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
\\
@merge @$m_1~m_2$ \> $=$ \> $@merge@'~(@init @m_1)~\{\}~(@init @m_2)~\{\}~@empty@$ \\
\\
$@merge@'$ \> $:$ \> $l_1 \times \{@Event@\} \times l_2 \times \{@Event@\}$               \\
           \> $\times$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
           \> $\to$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
\\
MMM \= M \= MMMMMMMMMM \= M \=\kill
$@merge@'~s_1~e_1~s_2~e_2~m$ \\
 \> $~|$ \> $(s_1, e_1, s_2, e_2) \in m$  \\
 \> $=$  \> $m$ \\
 \> $~|$ \> $(\mathit{type},\mathit{transitions}) \in @move@~s_1~e_1~s_2~e_2$ \\
 \> $=$  \> $\bigcup_{t~\in~\mathit{transitions}}(@merge@'~(@snd@~t)~m')$\\
~where  \> $m' = m~\cup~((s_1,e_1,s_2,e_2),~\mathit{type})~\cup~\mathit{transitions}$ \\
\end{tabbing}

@@

\subsubsection{Output and finishing}
The previous cases have been rather trivial as they require no synchronisation between the machines, but the remaining cases of @Out@, @OutFinished@, @Pull@, @Release@ and @Close@ are more complicated.

@@

The cases for @OutFinished@ proceed similarly to @Out@.

@@

If there is a @Finished@ in either machine's event set, and neither machine has a pending @Value@s to deal with, we know that the other machine has had either an @OutFinished@ or an empty @Pull@.
We proceed to the @None@ case of the @Pull@.

@@

\begin{code}
m1 : Pull xs     { Some => m2; None => m4 }
m2 : Out m (x+1) { Unit => m3 }
m3 : Release x   { Unit => m1 }
m4 : OutFinished m{ Unit => m5 }
m5 : Done
\end{code}

@@

\begin{code}
f1 : Pull xs     { Some => f2; None  => f5 }
f2 : If (x>0)    { True => f3; False => f4 }
f3 : Out f x     { Unit => f4 }
f4 : Release x   { Unit => f1 }
f5 : OutFinished f{ Unit => f6 }
f6 : Done
\end{code}

@@

\begin{code}
move m f m4 {Finished xs} f1 {Finished xs} =
    OutFinished m
    { Unit => m5, {Finished xs}, f1, {Finished xs} }
(Finish local output)
\end{code}

@@

Now @m@ is @Done@, and cannot move, while @f@ is at a @Pull xs@ that it already knows is finished.
\begin{code}
move m f m5 {Finished xs} f5 {Finished xs} =
    OutFinished f
    { Unit => m5, {Finished xs}, f6, {Finished xs} }
(Finish local output)
\end{code}

@@

The entire machine at the end is:
\begin{code}
m1,{},f1,{}
    : Pull xs
    { Some => m2, {Value xs},    f1, {Value xs}
    ; None => m4, {Finished xs}, f1, {Finished xs} }
m2,{Value xs},f1,{Value xs}
    : Out m (x+1)
    { Unit => m3, {Value xs},    f1, {Value xs}    }
m3,{Value xs},f1,{Value xs}
    : Skip
    { Unit => m1, {},            f1, {Value xs}    }
m1,{},f1,{Value xs}
    : Skip
    { Unit => m1, {},            f2, {Value xs}    }
m1,{},f2,{Value xs}
    : If (x>0)
    { True => m1, {},            f3, {Value xs}
    , False=> m1, {},            f4, {Value xs}    }
m1,{},f3,{Value xs}
    : Out f x
    { Unit => m1, {},            f4, {Value xs}    }
m1,{},f4,{Value xs}
    : Release x
    { Unit => m1, {},            f1, {}            }
m4,{Finished xs},f1,{Finished xs}
    : OutFinished m
    { Unit => m5, {Finished xs}, f1, {Finished xs} }
m5,{Finished xs},f1,{Finished xs}
    : Skip
    { Unit => m5, {Finished xs}, f5, {Finished xs} }
m5,{Finished xs},f5,{Finished xs}
    : OutFinished f
    { Unit => m5, {Finished xs}, f6, {Finished xs} }
m5,{Finished xs},f6,{Finished xs}
    : Done
\end{code}

@@

\begin{dot2tex}[scale=0.5]
digraph {
    rankdir=LR;
    mindist=0.1
    nodesep=0.1;
    ranksep=0.1;
    node [shape="circle", margin=0];
    10 [label="Pull xs"];
    20 [label="Out m (x+1)"];
    50 [label="If"];
    60 [label="Out f x"];
    70 [label="Release x"];
    900[label="OutFinished m"];
    920[label="OutFinished f"];
    930[label="Done"];

>>>>

\begin{tabbing}
MMMMMMM \= MM \= MMMMMM \= M\kill
@merge@ \> $:$ \> $@Machine @l_1 \times @Machine @l_2$ \\
        \> $\to$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
\\
@merge @$m_1~m_2$ \> $=$ \> $@merge@'~(@initial @m_1)~\{\}~(@initial @m_2)~\{\}~@empty@$ \\
\\
$@merge@'$ \> $:$ \> $l_1 \times \{@Event@\} \times l_2 \times \{@Event@\}$               \\
           \> $\times$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
           \> $\to$ \> $@Machine @(l_1 \times \{@Event@\} \times l_2 \times \{@Event@\})$ \\
\\
MMM \= M \= MMMMMMMMMM \= M \=\kill
$@merge@'~s_1~e_1~s_2~e_2~m$ \\
 \> $~|$ \> $(s_1, e_1, s_2, e_2) \in m$  \\
 \> $=$  \> $m$ \\
 \> $~|$ \> $(\mathit{type},\mathit{transitions}) \in @move@~s_1~e_1~s_2~e_2$ \\
 \> $=$  \> $\bigcup_{t~\in~\mathit{transitions}}(@merge@'~(@snd@~t)~m')$\\
~where  \> $m' = m~\cup~((s_1,e_1,s_2,e_2),~\mathit{type})~\cup~\mathit{transitions}$ \\
\end{tabbing}

@@

\subsubsection{Output and finishing}
The previous cases have been rather trivial as they require no synchronisation between the machines, but the remaining cases of @Out@, @OutDone@, @Pull@, @Release@ and @Close@ are more complicated.

@@

The cases for @OutDone@ proceed similarly to @Out@.

@@

If there is a @Finished@ in either machine's event set, and neither machine has a pending @Value@s to deal with, we know that the other machine has had either an @OutDone@ or an empty @Pull@.
We proceed to the @None@ case of the @Pull@.

@@

\begin{code}
m1 : Pull xs     { Some => m2; None => m4 }
m2 : Out m (x+1) { Unit => m3 }
m3 : Release x   { Unit => m1 }
m4 : OutDone m   { Unit => m5 }
m5 : Done
\end{code}

@@

\begin{code}
f1 : Pull xs     { Some => f2; None  => f5 }
f2 : If (x>0)    { True => f3; False => f4 }
f3 : Out f x     { Unit => f4 }
f4 : Release x   { Unit => f1 }
f5 : OutDone m   { Unit => f6 }
f6 : Done
\end{code}

@@

\begin{code}
move m f m4 {Finished xs} f1 {Finished xs} =
    OutDone m
    { Unit => m5, {Finished xs}, f1, {Finished xs} }
(Finish local output)
\end{code}

@@

Now @m@ is @Done@, and cannot move, while @f@ is at a @Pull xs@ that it already knows is finished.
\begin{code}
move m f m5 {Finished xs} f5 {Finished xs} =
    OutDone f
    { Unit => m5, {Finished xs}, f6, {Finished xs} }
(Finish local output)
\end{code}

@@

The entire machine at the end is:
\begin{code}
m1,{},f1,{}
    : Pull xs
    { Some => m2, {Value xs},    f1, {Value xs}
    ; None => m4, {Finished xs}, f1, {Finished xs} }
m2,{Value xs},f1,{Value xs}
    : Out m (x+1)
    { Unit => m3, {Value xs},    f1, {Value xs}    }
m3,{Value xs},f1,{Value xs}
    : Skip
    { Unit => m1, {},            f1, {Value xs}    }
m1,{},f1,{Value xs}
    : Skip
    { Unit => m1, {},            f2, {Value xs}    }
m1,{},f2,{Value xs}
    : If (x>0)
    { True => m1, {},            f3, {Value xs}
    , False=> m1, {},            f4, {Value xs}    }
m1,{},f3,{Value xs}
    : Out f x
    { Unit => m1, {},            f4, {Value xs}    }
m1,{},f4,{Value xs}
    : Release x
    { Unit => m1, {},            f1, {}            }
m4,{Finished xs},f1,{Finished xs}
    : OutDone m
    { Unit => m5, {Finished xs}, f1, {Finished xs} }
m5,{Finished xs},f1,{Finished xs}
    : Skip
    { Unit => m5, {Finished xs}, f5, {Finished xs} }
m5,{Finished xs},f5,{Finished xs}
    : OutDone f
    { Unit => m5, {Finished xs}, f6, {Finished xs} }
m5,{Finished xs},f6,{Finished xs}
    : Done
\end{code}

@@

\begin{dot2tex}[scale=0.5]
digraph {
    rankdir=LR;
    mindist=0.1
    nodesep=0.1;
    ranksep=0.1;
    node [shape="circle", margin=0];
    10 [label="Pull xs"];
    20 [label="Out m (x+1)"];
    50 [label="If"];
    60 [label="Out f x"];
    70 [label="Release x"];
    900[label="OutDone m"];
    920[label="OutDone f"];
    930[label="Done"];

### 2015-04-16 2015mergingmerges/section/05-Proofs.tex
change some @move@ rules over to new style

This proof proceeds very similarly to the above lemma.
For (Finish shared output, no pending events), a @Finished n@ is added to $e_2$, while @OutFinished n@ adds a @Finish n@ to the invariant set.
For (Shared pull, no pending values), a @Pull@ is emitted, and the @None@ case adds @Finished n@ to both sets, as with the invariant set.

@@

\paragraph{Lemma paths:} for two given states $u$ and $v$ of $@merge@~a~b$, and their corresponding input machine states $s$ and $t$, for each path between $u$ and $v$, there exists a path between $s$ and $t$ such that all states in the $st$ path are in the $uv$ path, \emph{except}:
\begin{itemize}
\item @Pull@s can be omitted if every predecessor path of the @Pull@ has a @Pull@, @Out@ or @OutFinished@ on the same channel;
\item @Release@s can be omitted if the channel is an output of the other machine;
\item @Release@s can be omitted if every successor path has a @Release@ on the same channel;
\item @Close@s can be omitted if every successor path has a @Close@ or @Pull@ on the same channel.
\end{itemize}

It is easy to see that for non-interfering transitions @Update@, @Skip@, @If@ and @Done@, this holds.
For the other cases, it is necessary to perform induction over the @move@ rules, and inspect the event sets.
If one of the input machine's states is a @Pull@, then either a @Pull@ or a @Skip@ may be produced.
@Skip@s are only produced if the event set already contains a @Value@ or a @Finished@, which can only be produced by an earlier @Pull@, @Out@ or @OutFinished@.
Likewise for the remaining cases.

>>>>

This proof proceeds very similarly to the above lemma.
For (Finish shared output, no pending events), a @Finished n@ is added to $e_2$, while @OutDone n@ adds a @Finish n@ to the invariant set.
For (Shared pull, no pending values), a @Pull@ is emitted, and the @None@ case adds @Finished n@ to both sets, as with the invariant set.

@@

\paragraph{Lemma paths:} for two given states $u$ and $v$ of $@merge@~a~b$, and their corresponding input machine states $s$ and $t$, for each path between $u$ and $v$, there exists a path between $s$ and $t$ such that all states in the $st$ path are in the $uv$ path, \emph{except}:
\begin{itemize}
\item @Pull@s can be omitted if every predecessor path of the @Pull@ has a @Pull@, @Out@ or @OutDone@ on the same channel;
\item @Release@s can be omitted if the channel is an output of the other machine;
\item @Release@s can be omitted if every successor path has a @Release@ on the same channel;
\item @Close@s can be omitted if every successor path has a @Close@ or @Pull@ on the same channel.
\end{itemize}

It is easy to see that for non-interfering transitions @Update@, @Skip@, @If@ and @Done@, this holds.
For the other cases, it is necessary to perform induction over the @move@ rules, and inspect the event sets.
If one of the input machine's states is a @Pull@, then either a @Pull@ or a @Skip@ may be produced.
@Skip@s are only produced if the event set already contains a @Value@ or a @Finished@, which can only be produced by an earlier @Pull@, @Out@ or @OutDone@.
Likewise for the remaining cases.
